---
title: "Documenting Changes in 01_download/README.Rmd"
author: "C.J. Tinant"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
    fig_caption: true
  github_document:
    toc: true
    toc_depth: 2
params:
  version: "v1.4.8"
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

Provide a human-readable summary of all download scripts in R/01_download/, including their purpose, input/output expectations, status, and key changes over time. This supports reproducibility and onboarding of new contributors.

## Overview
This section documents the current structure and status of all scripts responsible for downloading raw spatial and tabular data inputs. Each script is version-controlled and tied to a specific covariate or source dataset (See Table 1).

```{r table_of_scripts,message=FALSE,echo=FALSE,results='asis'}

library(knitr)
library(kableExtra)
library(tidyverse)

script_index <- tribble(
  ~`File Name`, ~Description, ~`Output Folder`, ~`Notes / Status`,
  "01a_download_ecoregions.R", "Download Level I–IV US ecoregion shapefiles", "data/raw/", "CRS standardized (EPSG:5070)",
  "01b_download_USGS_gage_data.R", "Download peak flow gage locations", "data/raw/", "Reorganized metadata",
  "01c_download_usgs_gage_metadata.R", "Query detailed site metadata", "data/raw/", "Outputs raw data-dictionary",
  "01d_download_usgs_peakflow_data.R", "Download annual peak flow values", "data/raw/", "Handles chunking & retries",
  "01e_filter_usgs_peakflow_data.R", "Filter data, calculate skew, tag PILFs", "data/processed/", "Replaces legacy script",
  "01f_download_nhdplus_v21.R", "Download NHDPlusV2.1 flowlines and catchments", "data/processed/", "CRS standardized (EPSG:5070)"
)

kbl(script_index, format = "latex", booktabs = TRUE, caption = "Script Index") %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  column_spec(2, width = "5.5cm") %>%
  column_spec(4, width = "5.5cm")

```

## Summary of Key Changes

### Standardized Folder Paths

- All output paths now use `here::here("data", "raw", ...)` or `data/processed/`
- Ensures reproducibility across systems and OSes

### Consistent CRS

- All spatial outputs reprojected to EPSG:5070 (CONUS Albers Equal Area)

### Modular Design

- Each script targets a single source or covariate
- Utility functions moved to `R/utils/`

### Logging & QA

- Retry attempts, download failures, and chunking behavior logged to `.csv` where applicable
- Scripts now include `{cli}` alerts for status tracking
- Each script is cross-linked to relevant milestone reports

---
