---
title: "Peak flow locs for Northwestern Great Plains and High Plains ecoregion"
author: "CJ Tinant"
date: "`r Sys.Date()`"
output: html_document
---

<!--
# Ecoregion hierarchy
Northwestern Great Plains and High Plains region of the SOUTH CENTRAL SEMI-ARID PRAIRIES of the GREAT PLAINS ecoregion

## Next steps
create an inline unzip
unzip /path/to/your/file.zip -d /path/to/destination/folder

-- need to add ecoreg when joining

* intersect level 3 ecoregion with station data
* sync personal and group Zotero library for FFA
https://guides.library.oregonstate.edu/c.php?g=359201&p=2426111
https://www.zotero.org/groups/5473862/olc_flood-frequency


## Helpful references:
GIT issues--
[GIT issues](https://dangitgit.com/)

Spatial data issues--
[Intro to spatial data in R](https://www.r4wrds.com/intro/m_intro_mapmaking)
[Overview of Coordinate Reference Systems (CRS) in R](https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf)
[Geocomputation with R](https://r.geocompx.org/)

References:
[dataRetrieval tutorial](https://waterdata.usgs.gov/blog/dataretrieval/)

[USGS watermapper](https://maps.waterdata.usgs.gov/mapper/index.html)
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# update sites_peak variable names -- to fit with l3 ecoregion names
sites_peak <- sites_peak %>%
  rename(na_l3_cde = ecoreg)

# check site codes -- zero obs in site_peak_vars_ck means all vars are described
sites_peak_vars <- tibble(code = names(sites_peak))

sites_peak_vars <- left_join(sites_peak_vars, site_criteria_cde,
                             by = join_by(code)) 

sites_peak_vars_ck <- sites_peak_vars %>%
  filter(is.na(description))

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
