---
title: "Documenting Changes in 02_clean_validate/README.Rmd"
author: "C.J. Tinant"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
    fig_caption: true
params:
  version: "v1.5.1"
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE,message=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(kableExtra)
library(tidyverse)
```

## Purpose

Provide a human-readable summary of scripts in `R/02_clean_validate`, including
their purpose, input/output expectations, status, and key changes over time.
This supports reproducibility and onboarding of new contributors.

## Overview
This section documents the current structure and status of scripts responsible
for processing files in `data/raw`. Each script is version-controlled and tied to a specific covariate or source dataset.

```{r table_of_scripts,results='asis'}

script_index <- tribble(
  ~`File Name`, ~Description, ~`Output Folder`, ~`Notes / Status`,
  "02a_merge_nhdplus_hr_flowlines.R", "Merge NHDPlus HR flowlines", "data/processed/nhdphr/", "CRS standardized (EPSG:5070)",
  "02b_merge_nhdplus_hr_catchments.R", "Merge NHDPlus HR catchments", "data/processed/nhdphr/", "CRS standardized (EPSG:5070)",
  "02c_validate_spatial_metadata.R", "Validate CRS and resolution of spatial files.", "data/log/", "",
  "02d_extract_metadata_from_script_names.R", "Generate a CSV tracker in covariate_source_inventory.Rmd", "data/log/", "Output is a table"
  )

kbl(script_index, format = "latex", booktabs = TRUE, caption = "Script Index") %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  column_spec(2, width = "5.5cm") %>%
  column_spec(4, width = "5.5cm")

```

## Summary of Key Changes

### Standardized Folder Paths

- Output paths outside of functions use `here::here("data", "processed")`

### Consistent CRS

- Spatial outputs reprojected to EPSG:5070 (CONUS Albers Equal Area)

### Modular Design

- Each script targets a single source or covariate
- Utility functions moved to `R/utils/`

### Logging & QA

- Retry attempts, download failures, and chunking behavior logged to `.csv` where applicable
- Scripts include `{cli}` alerts for status tracking

### Metadata Documentation

- Consistent headers at the top of each script.
- Script headers ready to be cross-linked to relevant milestone reports
