---
title: "Analysis Layout"
author: "CJ Tinant"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: true
---

```{r setup, echo=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.retina = 2)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggforce)

```

# Overview

The workflow is implemented in R using {tidyverse} and related packages to
reproducibly download, clean, join, and model covariate data for stream gage
locations.

### Primary objectives

1. **Model spatial variation in station skew** — the skew coefficient (Log-Pearson
Type III) using numeric covariates derived primarily from rasters aggregated by
Level II/III ecoregions and NHD+ catchments.

2. **Identify clusters of gaging stations** with similar covariate profiles to
support calculation of regional skew coefficients for the Great Plains ecoregion.

# Study Domain
The study focuses on USGS streamflow gaging stations in the **Great Plains ecoregion**.

- **Initial inventory:** All USGS streamflow gaging stations in the region
~11,000 stations.
- **Final analysis set:** ~1,700 unregulated stations with at least 20 years of
peak-flow data that meet Bulletin 17C quality criteria for skew estimation.

# Nested Scale Design Framework

Covariates span four domains:

- **Climate** (e.g., precipitation, temperature normals)
- **Land Cover** (e.g., cultivated fraction, vegetation types)
- **Topography** (e.g., elevation, slope, terrain curvature)
- **Watershed Metrics** (e.g., drainage area, shape indices)

To capture spatial heterogeneity and support robust regionalization, variables
are calculated at five spatial scales:

| Scale | Extent           | Sources                                       |
| :---: | :--------------- | :-------------------------------------------- |
| **0** | Station‑specific | Gage coordinates, elevation, drainage area    |
| **4** | Macroregional    | Aggregated EPA L3 to L4 Ecoregion boundaries  |
| **3** | Regional         | EPA L2 Ecoregion boundaries                   |
| **2** | Subregional      | EPA L3 Ecoregion boundaries                   |
| **1** | Local            | NHDPlusHD (1:24,000) catchment boundaries     |

## Variable counts by scale
```{r variable_counts_by_scale}

variable_count <- tibble::tribble(
  ~Scale, ~Extent, ~Climate, ~`Land Cover`, ~Topography, ~Watershed, ~Total,
  "0",  "Station-specific", 2, 0, 1, 1,  4,
  "4", "Macroregional"    , 3, 4, 3, 0, 10,
  "3", "Regional"         , 4, 4, 4, 5, 17,
  "2", "Subregional"      , 6, 2, 4, 3, 15,
  "1", "Local"            , 3, 2, 8, 4, 17
)

kable(variable_count, 
      caption = "Table 1. Covariate counts by scale and domain.") %>%
  kable_styling(full_width = FALSE)

```

# Covariate Descriptions
 
### Station-specific (Scale 0)
Station specific covariates are point-data at the location of the gage, and
include:

- **Longitude / Latitude** — spatial position, climate & physiographic gradients
- **Altitude** — influences temperature, snow persistence, runoff timing
- **Watershed Area** — key hydrologic scaling predictor

### Macroregional (Scale 4)
Macroregional-scale covariates are aggregated at the L3 to L4 Ecoregion levels
to create Tallgrass Prairie, Mixed-Grass Prairie, and Shortgrass-Steppe custom
macrozones, and include:

- **Climate Zone, PHZM Zone (Dominant)** — prevailing climate, cold-hardiness
- **PHZM Zone Count** — climate transition diversity
- **Land Cover Fractions** — cropland, forest, grassland, urban
- **Terrain Metrics** — mean & median slope, altitude zone

For additional details on aggregation, refer to:
`output/generalized_macroregions_map.png`, and
`output/final_macroregions_map.png`)

### Regional (Scale 3)
Regional scale covariates are aggregated at the L2 Ecoregion scale:

- **Annual Temperature / Precipitation** — baseline thermal & moisture regime
- **Pct May–Aug Precipitation** — storm seasonality
- **NDVI Metrics** — amplitude, IQR, peak NDVI, growing season length
- **Terrain Complexity** — slope mean/skewness/variability
- **Soil Texture Fractions** — clay, silt, sand
- **Stream Order** — median & max

### Subregional (Scale 2)
Regional scale covariates are data aggregated at the L3 Ecoregion scale:

- **Seasonal Precipitation Metrics** — fall, winter, spring, summer; stdev; IQR
- **MODIS Land Cover % / Diversity Index**
- **Soil Permeability, Runoff Class**
- **Topographic Wetness Index (TWI)** — mean, modal, class
- **Flow Accumulation** — drainage convergence

### Local (Scale 1)
Local or catchment scale covariates are aggregated at a NHD+ HD (1:24,000)
resolution level, and include:

- **Climatic Intensity** — freeze–thaw days, Precipitation intensity, wet-day frequency
- **NLCD Land Cover % / Diversity Index**
- **Terrain Morphology** — elevation range, aspect, curvature, relief ratio
- **Watershed Geometry** — elongation ratio, circularity ratio
- **Drainage Network** — stream density, flow length, stream slope

# Reference Year
All covariates use 2016 as the reference year because it lies near the midpoint
of the 1991–2020 normals, is broadly available at high quality (PRISM, MODIS,
NLCD), and avoids major anomalies—supporting consistent, comparable modeling
across datasets and scales.

# Supporting Scripts

`R/01_download/01b_download_USGS_gage_data.R` -- Download & filter gage sites for
Great Plains.

`R/01_download/01d_download_usgs_peakflow_data.R` -- Download annual peak‑flow
values with batching/retry.

`R/03c_make_macrozone_lut.R` and `R\03d_make_macrozone_layer.R` -- Delineate
L2 to L4 Ecoregions 

See `docs/download_scripts.md` for the full script index and status.

