---
title: "Check Raw Data Files"
author: "CJ Tinant"
output:
  github_document:
    toc: true
    toc_depth: 3
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
---

## Overview

This document supports QA of raw spatial inputs used for regional skew modeling.
The goal is to inventory, validate, and deduplicate vector and raster files
within the data/raw/ directory prior to covariate extraction.

## Goals

-   ✅ Identify and remove duplicate files from data/raw
-   ✅ Deduplicate and standardize raster inputs in data/raw/
-   🗂️ Organize spatial datasets by relevance, format, and source
-   📝 Document retained files in data/meta/spatial_source_checklist.csv

## Workflow Part 1: Deduplication of Vector Spatial Files

### 1. Inventory Vector Files

- List vector files recursively and flag potential duplicates

### 2. Summarize by Name and Size

- Group by basename
- Highlight different file sizes as potential version conflicts
- Flag preferred formats (e.g., keep .gpkg, review .shp)

### 3. Save Results to CSV

**Exported outputs:**
- to_check/vector_file_inventory.csv – full file list with duplicate flags
- to_check/duplicate_vector_summary.csv – filtered table of same-name files

### 4. Draft Cleanup Plan

- Move lower-priority duplicates to to_check/duplicates/
- Confirm projections and attribute consistency before deletion

### 5. Documentation

- Add exploratory summaries and notes on retained files
- Optionally create data/meta/spatial_source_checklist.csv
- Tag major cleanup step as v0.3.1-dedup-raw-vector
- Append to milestone_02_documentation.Rmd

```{r setup_raster, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)              # simple features in R
library(fs)              # file system operations
library(here)            # find files
library(tidyverse)

file.exists(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))
file.exists(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))

source(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))
source(here("R/utils/clean_organize_spatial_data/move_low_priority_duplicates.R"))

```

### Run Vector Deduplication Inventory Function

```{r run-dedup-vector-function}

result <- dedup_vector_inventory()

```

### Move Zipped Files to Archives

```{r run-zipped-to-archives-function}

source(here("R/utils/clean_organize_spatial_data/move_zip_to_archives.R"))
move_zip_to_archives()

```

## Workflow Part 2: Deduplication of Vector Spatial Files

### 1. Inventory Raster Files

- Search .bil, .tif, .img, .zip, .hdr, .stx

- Flag files with same name and different sizes

- Group by basename and detect variations

### 2. Save Results to CSV

**Exported outputs:**
- to_check/raster_file_inventory.csv – full file list with duplicate flags
- to_check/duplicate_vector_summary.csv – filtered table of same-name files

### 3. Draft Cleanup Plan

- Move lower-priority duplicates to to_check/duplicates/
- Confirm projections and attribute consistency before deletion

### 4. Documentation

- Add exploratory summaries and notes on retained files
- Optionally create data/meta/spatial_source_checklist.csv
- Tag major cleanup step as v0.3.1-dedup-raw-vector
- Append to milestone_02_documentation.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)              # simple features in R
library(fs)              # file system operations
library(here)            # find files
library(tidyverse)

file.exists(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))
file.exists(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))

source(here("R/utils/clean_organize_spatial_data/dedup_vector_inventory.R"))
source(here("R/utils/clean_organize_spatial_data/move_low_priority_duplicates.R"))

```

### Run Raster Deduplication Inventory Function

```{r run-dedup-raster-function}

source(here("R/utils/clean_organize_spatial_data/dedup_raster_inventory.R"))

result_ras <- dedup_raster_inventory()

```
